---
title: "11082245_BIOL63101_Assignment2"
author: '11082245'
date: "2022-12-15"
output: 
  rmdformats::downcute:
    self_contained: true
    thumbnails: false
    lightbox: true
    gallery: false
    toc_float: TRUE
    default_style: "dark"
    fig_caption: true
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(cache = TRUE)
```

# Before Reading

The R script, R Markdown and this `.html` file were uploaded to [my GitHub repository](https://github.com/azanghai/UoM_BIOL63101_Assignment_2), where you can view or download my code. I have also uploaded the .html file to my website. You can click [here](https://wentingxu.netlify.app/assignment/11082245_biol63101_assignment2.html) to view. In order to make the code look neater, the code already explained in the text will not be explained again.

**To get the best browsing experience, maximize the browser.**

**You can use the switch button on the left if you want to switch to a different colour theme.**

------------------------------------------------------------------------

# Before start

Before starting the analysis, we want to load our packages first.

```{r, import_libraty_chunk, message=FALSE, warning=FALSE}
# load tidyverse for reading, wrangling and plotting data
library(tidyverse) 
# load afex for ANOVA functions
library(afex) 
# load emmeans for pairwise comparisons
library(emmeans) 
# load psych for data summarize 
library(psych) 
# load skimr package for data summarize
library(skimr)
# load ggpubr package for plotting
library(ggpubr)
# load car package for leveneTest
library(car)
```

# Question 1

## Load Data and Preliminary Processing

Here, by using the function `read_csv()` in the `tidyverse` package, we can load `.csv` file as a data frame in RStudio.

```{r, message=FALSE, warning=FALSE}
#load data
Q1_data = read_csv('./Raw_Data/assignment_2_dataset_1.csv')

# check if the data is correctly loaded
head(Q1_data)
```

We find that: \* the value from condition needs to be renamed for better understanding. \* Column `condition` needs to be converted to factors.

```{r}
# rename the value from condition_a, condition_b to normally and degraded.
# here we use mutate() function to recreate the column.
# case_when() is to select the case that fits different "condition".
Q1_data = Q1_data %>%
    mutate(
        condition = case_when(
            condition == "condition_a" ~ "normally",
            condition == "condition_b" ~ "degraded"
        )
    )

# code condition as a factor,using mutate function and factor function to replace the old column.
Q1_data = Q1_data %>%
    mutate(condition = factor(condition))

# double check if the data is coded correctly.
head(Q1_data)
```

Now the data looks correctly loaded.

## Summarize the Data {.tabset .tabset-fade .tabset-pills}

Here I have used different measures to summarize the data, the result is the same, but the method provides a different aspect of the data.

### Using Base R Functions

We can use base R functions to summarize the data by using the following code.

```{r}
# summarize data using base R functions
Q1_data %>%
    # we want to view the data in different conditions.
    group_by(condition) %>%
    # here we calculate mean and sd for the data, using summarize()
    summarise(mean = mean(response_time),
              sd = sd(response_time))
```

### Using `psych` Package

`psych` is a package coded explicitly for psychology students. It has functions for personality, psychometric theory and experimental psychology purposes. Here we use the `describeBy()` function to summarize the data. I use this package because it provides values from the number of valid cases to the skew, which gives us a good understanding of the data.

```{r}
# according to this package's documentation, we use `describeBy()` function to compare data in different groups.
# The first argument sent is the data frame, and the second one is the group list.
describeBy(Q1_data$response_time, Q1_data$condition)
```

### Using `skimr` Package

`skimr` is also a package that helps us summarize the data. I use this package because it can compare the data in one chart. And it's also easy to use.

```{r}
Q1_data %>%
    group_by(condition) %>%
    # here we use skim() function to generate summarize.
    skim() %>%
    # as we are only interested in the response_time variable, we use filter() to select it and compare between groups.
    filter(skim_variable == "response_time")
```

## Data Visualization

Here, by visualizing the data, we can have a first impression of whether there are differences in the data.

```{r, message=FALSE, warning=FALSE,fig.align='left',out.width='100%',out.height='100%'}
# Seed is set here to ensure reproducibility.
set.seed(11082245)
# use ggplot to demonstrate the distribution of data.
Q1_data %>%
    #here we set x axis as condition, y axis as response time, and we coloured the data according to the different conditions.
    ggplot(aes(x = condition, y = response_time, colour = condition)) +
    # here we add a layer of violin plot using settings and parameters in ggplot()
    geom_violin() +
    # let the dots jitter a little bit to have a better look.
    geom_jitter(width = .1) +
    # the figure is clear enough and we do not need a guides.
    guides(colour = FALSE) +
    # here we use stat_summary() function to add another layer.
    # "mean_cl_boot" is a built in function to acquire confidence limits for the mean.
    stat_summary(fun.data = "mean_cl_boot", colour = "black") +
    # also, a theme and text size is set here for a better view.
    theme_minimal() +
    theme(text = element_text(size = 13)) 
```

By viewing the plot, we have also found no abnormal value. So we don't need to remove the abnormal value

## Normality and Homogeneity of Variance Test

As there are three assumptions in ANOVA.\
- The responses for each factor level have a normal population distribution.\
- These distributions have the same variance.\
- The data are independent.

Though the program will calculate and correct the variance, and the central limit theorem tells us that no matter what distribution things have, the sampling distribution tends to be normal if the sample is large enough (n \> 30). Still, we would like to perform a test for Normal Distribution and Homogeneity of Variance

### Normality Test {.tabset .tabset-fade .tabset-pills}

#### Q-Q plot

We can draw a Q-Q plot to find whether the data is distributed normally.

As we can see, if the dots are distributed around the line `y = x`, it means that the distribution of the data is likely to be Normal distribution. But to be certain, we also need to perform a Normal Distribution Test

```{r,fig.align='left',out.width='100%',out.height='100%'}
ggplot(data = Q1_data, aes(sample = response_time, colour = factor(condition))) +
    geom_qq() +
    geom_qq_line() +
    facet_wrap( ~ condition) +
    guides(colour = FALSE) +
    labs(y = "Response Time", x = "Normal Theoretical Quantiles") +
    theme_minimal() +
    theme(text = element_text(size = 13))
```

From the plot, we can see that the dots for different data group is closely attached to the `y = x` line. Therefore, the data may be normally distributed.

#### Shapiro-Wilk Normality Test

Then we perform a Shapiro-Wilk normality test.

Here we use the `tapply()` function. The first argument is the data frame, and the second is the Classification parameters. The third one is the function that we want to perform. This function allows us to perform repeated functions using one line of code.

```{r}
tapply(Q1_data$response_time, Q1_data$condition, shapiro.test)
```

For the output, the P-value \> 0.05. Therefore we could say that the distribution of the data is not significantly different from the normal distribution. In other words, we can assume the normality.

### Homogeneity of Variance Test

Here we perform the Levene's Test for Homogeneity of Variance only because this is one of the most commonly used methods.

This function is in the `car` package. Just like the liner model, we load the factor and the variable in the function to get a result.
```{r}
leveneTest(response_time ~ condition, data = Q1_data)
```

As the p-value is higher than 0.05(0.5175), Therefore we accept the null hypothesis (H0): the variances of the two groups are equal.

## Building ANOVA model

Here, we build the model by considering `response_time` as DV, and `condition` as IV. Meanwhile, `participant` is considered as the source of random error.

```{r, message=FALSE, warning=FALSE}
Q1_model_1 = aov_4(response_time ~ condition + (1 | participant),data = Q1_data)
```

Then, we would like to take a look at the model we built and perform a post-hoc analysis.

```{r}
summary(Q1_model_1)
# by using emmeans(), we could compare the data group by the condition factor.
emmeans(Q1_model_1, pairwise ~ condition)
```

We can also view the result in a plot by adding several layers on the plot we have made. Function `stat_compare_means()` is used to accomplish this task. By setting `comparisons` to `degraded` and `normally`, we draw a line to indicate the p-value between these groups. Also, by setting `symnum.args`, we can convert the p-value to symbols for a better understanding.

```{r,fig.cap='***p<0.001',warning=FALSE,message = FALSE,results='hide',fig.align='left',,out.width='100%',out.height='100%'}
Q1_data %>%
    ggplot(aes(x = condition, y = response_time, colour = condition)) +
    geom_violin() +
    geom_jitter(width = .1) +
    # here we add ANOVA result to the plot
    # note this anova function may not get the same result.
    stat_compare_means(method = "anova") +
    # here we draw the line to compare different groups
    stat_compare_means(comparisons = list(c("degraded", "normally")),
                       symnum.args = list(
                           # use * to illustrate the p value
                           cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf),
                           symbols = c("****", "***", "**", "*", "ns")
                       )) +
    guides(colour = FALSE) +
    theme_minimal() +
    theme(text = element_text(size = 13)) +
    labs(y = "Response Time", x = "Conditions")
```

### T-test

A t-test is available for the data analysis as the experiment is a between-participants design and only has a factor with two levels. 

By using the built-in `t.test()` function, the first parameter we sent is the data frame, and the second one is the group value. The function will perform a t-test based on the parameter we sent.

```{r}
t.test(Q1_data$response_time ~ Q1_data$condition)
```

The result is similar, the difference is significant.

## Conclusion

From the analysis performed. It can be inferred that the difference between response time caused by different visual quality is significant. This means the subject will respond quicker(estimated marginal means = 1002) with normally present words than visually degraded(estimated marginal means = 1020).

# Question 2

## Load Data and Preliminary Processing

Similar to Question 1, we want to load the data and perform some preliminary processing.

```{r, message=FALSE, warning=FALSE}
Q2_data = read_csv('./Raw_Data/assignment_2_dataset_2.csv')
head(Q2_data)

Q2_data = Q2_data %>%
    mutate(
        condition = case_when(
            condition == "condition_a" ~ "normally",
            condition == "condition_b" ~ "degraded"
        )
    )
head(Q2_data)
```

Also, we want to convert `condition` in to factor.

```{r}
Q2_data = Q2_data %>%
    mutate(condition = factor(condition))
head(Q2_data)
```

## Summarize the Data

Here, we use `skimr` package to summarize the data. In this case, we want to learn the distribution of the data under the condition `caffeine` and `condition`.

```{r}
Q2_data %>%
    group_by(condition) %>%
    skim() %>%
    filter(skim_variable != "participant")

Q2_data %>%
    group_by(caffeine) %>%
    skim() %>%
    filter(skim_variable == "response_time")
```

However, this is still too complicated. Data visualization is considered to demonstrate the data in a more efficient way.

## Data Visualization

Here, we built a scatter plot. By viewing the plot, it seems that there is a relationship between response time and caffeine consumption.

```{r,fig.align='left',out.width='100%',out.height='100%'}
# we set colour = condition in order to distinguish the relation between caffeine consumption and conditions. 
ggplot(Q2_data, aes(x = caffeine, y = response_time, colour = condition)) +
    geom_point(size = 3, alpha = .9) +
    labs(x = "Caffeine Consumption (cups of coffee)",
         y = "Response Time") +
    theme_minimal() +
    theme(text = element_text(size = 11)) 
```

Again, as the data is identical to question 1, we do not need to repeat the normality and homogeneity of variance test, nor do we need to remove any abnormal data.

## Building Model {.tabset .tabset-fade .tabset-pills}

### ANOVA Model

In this case, we set `response` as DV and condition as IV. Also, we add caffeine as a covariate. Meanwhile, we set `factorize` as false to avoid the program reading `caffeine` as a factor.

```{r, message=FALSE, warning=FALSE}
Q2_model = aov_4(response_time ~ caffeine + condition +(1 | participant),data = Q2_data,factorize = FALSE)
```

After that, we use the `anova()` function to acquire the effect size. And using `emmeans()` to compare different groups.

```{r}
anova(Q2_model)
# here we use 'pairwise' to compare different conditions
emmeans(Q2_model, pairwise ~ condition)
```

From the result, it seems that the effect of covariate and conditions is not significant in this case. And the adjusted means are slightly different when considering the influence of caffeine(normally from 1002 to 1005, degraded from 1020 to 1018).

### linear model

Here, we use the `lm()` function to build a linear model. But first, we want to set the factor level.

```{r}
Q2_data <- Q2_data %>%
    mutate(condition = fct_relevel(condition,
                                   c("normally", "degraded")))
# here we use contrasts() to check the contrasts
contrasts(Q2_data$condition)
```

We are doing this because this will set `normally` as the intercept.

```{r}
Q2_model_lm = lm(response_time ~ condition + caffeine, data = Q2_data)
Q2_model_lm
```

At this step we can finally write the function as `Resopnse Time = 998.564 + 12.791*degraded + 2.469*mean(Q2_data$caffeine)`. Therefor we can compare the result with ANOVA result.

```{r}
# for ANOVA result: normally = 1005
Resopnse_Time_normally = 998.564 + 12.791 * 0 + 2.469 * mean(Q2_data$caffeine)
Resopnse_Time_normally

# for ANOVA result: degraded = 1018
Resopnse_Time_degraded = 998.564 + 12.791 * 1 + 2.469 * mean(Q2_data$caffeine)
Resopnse_Time_degraded
```

As we can see, the result is rounded to the ANOVA result. That means they have the same result.

However, it seems to me that there is no clear "order" between the two levels under the conditions. They are just different categories. So it is a little bit "weird" to use "normally" as an intercept. To solve this feeling, I find a way to set the intercept to zero.

```{r}
Q2_model_lm_1 = lm(response_time ~ condition - 1 + caffeine, data = Q2_data)
Q2_model_lm_1
```

I find this method on [stack overflow](https://stackoverflow.com/questions/41032858/lm-summary-not-display-all-factor-levels), It seems that the `-1` change the order of the factor level.

Also, there is another way to add a `0` in the formula. This sets the intercept to zero.

```{r}
Q2_model_lm_2 = lm(response_time ~ 0 + condition + caffeine, data = Q2_data)
Q2_model_lm_2
```

The function will be `Response Time = 998.564*normally + 1011.354*degraded + 2.469*mean(Q2_data$caffeine)`. But still we will get the same result compared to ANOVA data.

```{r}
# for ANOVA result: normally = 1005
Response_Time_0_1 = 998.564 * 1 + 1011.354 * 0 + 2.469 * mean(Q2_data$caffeine)
Response_Time_0_1

# for ANOVA result: degraded = 1018
Response_Time_0_2 = 998.564 * 0 + 1011.354 * 1 + 2.469 * mean(Q2_data$caffeine)
Response_Time_0_2
```

# Question 3

## Load Data and Preliminary Processing

```{r, message=FALSE, warning=FALSE}
Q3_data_raw = read_csv('./Raw_Data/assignment_2_dataset_3.csv')
head(Q3_data_raw)
```

Here, we find that we need to convert data frame to longer format.

```{r}
Q3_data_converted <- Q3_data_raw %>%
    # pivot function can convert wider data frame to longer data frame
    pivot_longer(
        # here we set columns that need to be converted
        cols = c(
            positiveprime_positivetarget,
            positiveprime_negativetarget,
            negativeprime_positivetarget,
            negativeprime_negativetarget
        ),
        # we send two column name here to indicate the function to convert them into two columns
        names_to = c("prime", "target"),
        # here we tell the function to separate conditions at the '_' symbol
        names_sep = "_" ,
        # here we tell the program where to save the value
        values_to = "rt",
        # set the transformed variable to be factor
        names_transform = "factor"
    )

# manually check the data
head(Q3_data_converted)
```

## Summarize the Data

Here, we use `skimr` package to summarize our data.

```{r}
Q3_data_converted %>%
    # in this case, we use two grouping variables
    group_by(prime, target) %>%
    skim() %>%
    filter(skim_variable == "rt")
```

From the result, we can find that there is no missing data, and it seems that there is a difference between coherent and incoherent prime-target pairs.

## Data Visualization

```{r,,fig.align='left',out.width='100%',out.height='100%'}
Q3_data_converted %>%
    # here we set x axis to be prime and target pair, which helps us compare the data
    ggplot(aes(
        x = prime:target,
        y = rt,
        colour = prime:target
    )) +
    geom_violin() +
    geom_jitter(width = .1, alpha = .25) +
    guides(colour = FALSE) +
    stat_summary(fun.data = "mean_cl_boot", colour = "black") +
    theme_minimal() +
    # because the tag is too long, we need to rotate it a little bit.
    theme(text = element_text(size = 13),
          axis.text.x = element_text(
              angle = 35,
              vjust = 0.5,
              hjust = 0.5
          )) +
    labs(x = "Prime X Target", y = "RT")
```

## Building Model

As the data are from a 2 x 2 repeated measures design. Therefore we use `prime * target` to be the predictor. This corresponds to two main effects, plus the interaction between them. Meanwhile, because it is a repeated measures design, we want to add `prime * target` to the random effect.

```{r, message=FALSE, warning=FALSE}
Q3_model = aov_4(rt ~ prime * target + (1 + prime * target | participant), data = Q3_data_converted)
Q3_model
```

And we use the `anova()` function to get a more detailed result. Meanwhile, use `emmeans()` to compare different levels. I have chosen the Bonferroni method because ~~it seems that the Tukey method is not suitable for repeated measures design~~(I've seen it somewhere on stack overflow, and it is recommended in our course). 

```{r}
# we use this function to get a more specific result
anova(Q3_model)

# here we use
emmeans(Q3_model, pairwise ~ prime * target, adjust = "Bonferroni")
```

## Conclusion

According to the result, we can infer that the difference between `negativeprime negativetarget - positiveprime negativetarget` is statistically significant(1547 to 1567, t(294)=3.138, *p* = 0.0123). And the difference between `negativeprime positivetarget - positiveprime positivetarget` is also statistically significant(1563 to 1547, t(294)=2.908, *p* = 0.0252). It means that people could responded faster to positive images following a positive prime, and also can responded faster to negative images following a negative prime. Meanwhile, We conducted a 2 (Prime: Positive vs. Negative) x 2 (Target: Positive vs. Negative) repeated measures ANOVA to investigate the influence of prime valence on reaction times to pictures of Positive or Negative valence. The ANOVA revealed no effect of prime (F(1,147) = .3157, *p* = .5751, ηG2 = .0004908), no effect of target (F(1, 147) = .2375, *p* = .6267, ηG2 = .0004315), but an interaction between target and prime (F(1, 147) = 17.2541, p = <.001, ηG2 = .0292770).

For a easier way to understand this, I have made a plot. 

```{r,fig.cap='*p<0.05, **p<0.01', message=FALSE, warning=FALSE,fig.align='left',out.width='100%',out.height='100%',results='hide'}
Q3_data_converted %>%
    # here we set x axis to be prime and target pair, which helps us compare data
    ggplot(aes(
        x = prime:target,
        y = rt,
        colour = prime:target
    )) +
    geom_violin() +
    geom_jitter(width = .1, alpha = .25) +
    guides(colour = FALSE) +
    stat_summary(fun.data = "mean_cl_boot", colour = "black") +
    theme_minimal() +
    # because the tag is too long, we need to rotate it a little bit.
    theme(text = element_text(size = 13),
          axis.text.x = element_text(
              angle = 35,
              vjust = 0.5,
              hjust = 0.5
          )) +
    labs(x = "Prime X Target", y = "RT") +
    # here we draw the line to compare groups
    stat_compare_means(comparisons = list(
        c(
            "negativeprime:negativetarget",
            "positiveprime:negativetarget"
        ),
        c(
            "positiveprime:positivetarget",
            "negativeprime:positivetarget"
        )
    ),
    symnum.args = list(
        # use * to illustrate the p value
        cutpoints = c(0, 0.0001, 0.001, 0.01, 0.05, Inf),
        symbols = c("****", "***", "**", "*", "ns")
    ))
```

<center>------✨🎄Merry Christmas🎄✨------</center>
